% -*- mode:LaTex; mode:visual-line; mode:flyspell; fill-column:75-*-

% Special indentation for acknowledgments.
\setlength{\parskip}{1em}
\setlength{\parindent}{0em}

\noindent
First, I am grateful to Prof. Michael Kaess, my advisor for his patience, guidance and faith in me over the past two years through thick and thin! Without his support and ideas through our discussions, notwithstanding funding, this work would not have been possible. I would like to thank Prof. Katerina Fragkiadaki, for her guidance through the third semester about deep learning, and impressing the need to stay up-to-date  with the research literature, and providing avenues to hone my learning. I am thankful to Prof. Sebastian Scherer for serving on my committee on short notice, and for asking insightful and incisive questions! I acknowledge Chaoyang Wang for serving on my committee and for interesting discussions on future work.

I would like to thank Wei Dong for being foremost a friend, a supportive co-author and for the insightful research discussions, that significantly influenced this work. I am thankful to my colleagues and former labmates in the Robot Perception lab (RPL) for their encouragement, honest research feedback and comments. I have learned a lot from them, and they made work at CMU an enjoyable experience.

I would like to thank some of my friends and colleagues in the Master of Science in Robotics cohort of '21, in particular Tarasha Khurana for the support and riveting discussions that ensued about research. Last, but not least, I'd like to thank my parents and family for their faith and support.
